---

output: github_document
---

Assignemnt 1
=================================

```{r setoptions, echo=FALSE}
library('knitr')
opts_chunk$set(echo=TRUE, keep_md = TRUE)
```

Let us read the data. THe following code assumes that the csv file is already loaded and unzipped into the working directory. 

```{r}
dat<-read.csv("activity.csv")
```
Let us calculate the total number of step per day. The Na's are just ignored. 
After calculation the histogram is build, which shows the correspoding distribution density.
```{r}
total_day<-aggregate( steps ~ date, data = dat, FUN = sum, na.rm = TRUE)
head(total_day)
hist(total_day$steps, xlab="steps pers day", ylab=" Counts", col="blue", breaks=20)
```

Calculate the mean and median number of steps taken per day. 
```{r}
steps_mean<- mean(total_day$steps)
steps_mean
steps_median<- median(total_day$steps)
steps_median
```
##What is mean total number of steps taken per day?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
total_interval<-aggregate( steps ~ interval, data = dat, FUN = mean, na.rm = TRUE)
plot(total_interval$interval, total_interval$steps, type="l", xlab="5-min interval", 
             ylab= "Total steps", col="blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Max number of steps
```{r}
max_steps_int<-max(total_interval$steps)
max_steps_int
```
**Interval containing max number of steps in average:**
```{r}
#interval with max number of steps
total_interval$interval[which(total_interval$steps==max_steps_int)]
```
##Imputing missing values


How many missing values do we have?
```{r}
steps_na<-length(dat$steps[dat$steps=="NA"])
steps_na
```

looks like a lot!!!!

How does it look like?
```{r}
head(dat$steps, 20)
```

###Let us impute the missing values with average numer of steps per corresponding tiem interval. 

Calculating mean number of steps per interval, ignoiring missing values. Showing first 30 intervals. 

```{r}
mean_interval<-aggregate( steps ~ interval, data = dat, FUN = mean, na.rm = TRUE)
head(mean_interval, 30)
```
Creating an impute function, which replaces the missing values with the mean over the interval. The later are looked up in the dataframe *mean_interval*
```{r}
impute<-function(interval){
        i<-which(mean_interval$interval==interval) 
        mean_interval$steps[i]
}
```
Let us now replace all the NAs with the corresponding means. 
```{r}
dat$steps[is.na(dat$steps)]<-impute(dat$interval[is.na(dat$steps)])
head(dat$steps, 20)
```
It looks like we've done that!!!

Now we make the same plots but with the rectified dataset. 
Total number of steps per day:
```{r}
total_day_imp<-aggregate(steps ~ date, data = dat, FUN = sum, na.rm = TRUE)
hist(total_day_imp$steps, xlab="steps pers day", ylab=" Counts", col="blue", breaks=20)
```

Calculate the mean and median number of steps taken per day. 
```{r}
steps_mean_imp<- mean(total_day_imp$steps)
print("mean od number of steps per day")
steps_mean_imp
steps_med_imp<- median(total_day_imp$steps)
print("meadian of number of steps per day")
steps_med_imp
```

**There is not much difference in this particular case between original and imputted datasets**


##Are there differences in activity patterns between weekdays and weekends?


is it a weekday or weekend? creating a function, which returns the answer:
```{r}
wd<-function(date){
        if(weekdays(date)=="Saturday" | weekdays(date)=="Sunday") as.factor("weekend")
        else as.factor("weekday")
}
```

Creating a corresponding factor variable in the imputed dataset. 
```{r}
dat$wdays<-sapply(as.Date(dat$date), wd)
head(dat, 10)
```

calculating mean number of steps per interval with the rectified dataset. Showing first 20 intervals. 

```{r}
mean_interval_imp<-aggregate( steps ~ interval + wdays, data = dat, FUN = mean, na.rm = TRUE)
head(mean_interval_imp, 20)
```

Make a panel plot containing a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
library(ggplot2)
ggplot(mean_interval_imp, aes(x =interval , y=steps, color=wdays)) +
       geom_line() +
       labs(title = "Mean Daily Steps (type of day)", x = "Interval", y = "Total Number of Steps") +
       facet_wrap(~ wdays, ncol = 1, nrow=2)
```
